Screening Sinkhorn Algorithm for Regularized Optimal Transport

### 2019-05-20 tex/v3/section_5_numeric_expes.tex
change intro
mes commit

We compare \emph{Screenkhorn} to Sinkhorn as implemented in Python Optimal Transport toolbox\footnote{see \url{https://pot.readthedocs.io/en/stable/index.html}} on  a synthetic example. The dataset we use consists of source samples generated from a bi-dimensional gaussian mixture and target samples following the same distribution but with different gaussian means. We consider an unsupervised domain adaptation using optimal transport with entropic regularization.  Several settings are explored: different values of $\eta$, the regularization parameter, the allowed budget $\frac{n_b}{n} = \frac{m_b}{m} = $ ranging from $0.01$ to $0.99$, different values of $n$ and $m$ and finally, whether the cost matrix $C$ is normalized or not.

>>>>

We compare \emph{Screenkhorn} to Sinkhorn as implemented in Python Optimal Transport toolbox\footnote{\url{https://pot.readthedocs.io/en/stable/index.html}} on  a synthetic example. The dataset we use consists of source samples generated from a bi-dimensional gaussian mixture and target samples following the same distribution but with different gaussian means. We consider an unsupervised domain adaptation using optimal transport with entropic regularization.  Several settings are explored: different values of $\eta$, the regularization parameter, the allowed budget $\frac{n_b}{n} = \frac{m_b}{m} = $ ranging from $0.01$ to $0.99$, different values of $n$ and $m$ and, whether the cost matrix $C$ is normalized or not. We empirically measure  marginal violations $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ as well as running time expressed as $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$
Figure \ref{fig:margin_expe} summarizes the observed behaviors of both algorithms under these settings. We choose to only report results for $n=m=1000$ as we get similar findings for other values of $n$ and $m$.

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=14.cm]{../../figure/norm_M_marginals_toy_n1000} \\
		\includegraphics[width=8.cm]{../../figure/norm_M_time_toy_n1000}
	\end{center}
	\caption{Empirical evaluation of \emph{Screenkhorn} vs Sinkhorn for normalized cost matrix. (Top panel) marginal violations in relation with the budget of points and (Bottom panel) ratio of computation times $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$. The results are averaged over 10 runs.} 
		\label{fig:margin_expe}
\end{figure*}

\emph{Screenkhorn} provides good approximation of the marginals $\mu$ and $\nu$ for ''high'' values of the regularization parameter $\eta$ ($\eta > 1$). The approximation quality diminishes for small $\eta$. As expected $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ converge to zero when increasing the budget of points. Remarkably those quantities are all almost null whatever the budget for high $\eta$.  According to computational gain, \emph{Screenkhorn} is, up to 2 times; faster than Sinkhorn at low budget $n_b/n$
